---
title: "RepData_Week2 Assignment Peer-graded Assignment: Course Project 1"
author: "Daisy Niu"
date: "10/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Load library needed for data analysis/plotting
```{r}
library(ggplot2)
library(dplyr)
```

## Set working directory and Download/unzip data to the working directory.
```{r}
setwd("C:/OnlineCourse/DataScience/ReproductibleResearch/Week2_Project")

download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip","repdata_data_activity.zip")

unzip("repdata_data_activity.zip")

```
## Read data and pre-process data for analysis

```{r}

Activity<-read.csv("activity.csv")
head(Activity)

```


## What is mean total number of steps taken per day?
### 1. Calculate the total number of steps taken per day
```{r}

Step_Date<-aggregate(steps~date, Activity, sum, na.rm=TRUE)
head(Step_Date)

```
### 2. Histogram of the total number of steps taken per day
```{r}
qplot(steps,data=Step_Date,xlab='Total steps per day', ylab='Frequency using binwith 500',main='Histogram of total steps taken each day', binwidth=500)
```

### 3. Calculate mean and median of the total number of steps taken per day
```{r}
mean(Step_Date$steps)
median(Step_Date$steps)

```

## What is the average daily activity pattern?
### 1. Make a time series plot
```{r}
Step_interval<-aggregate(steps~interval, Activity, mean, na.rm=TRUE)
head(Step_interval)

qplot(interval,steps, data=Step_interval,geom='line', xlab = '5-minute intervals',ylab = 'Average steps taken across all days',color="red")
```

### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
Step_interval[which.max(Step_interval$steps),]
```

## Imputing missing values
### 1. Calculate and report the total number of missing values in the dataset 
```{r}
NA_Activity<-Activity[which(is.na(Activity$steps)),]
head(NA_Activity)
nrow(NA_Activity)

```

### 2-3. Create a new dataset that is equal to the original dataset but with the missing data filled in. Strategy: Replace NA with the mean steps of each interval.
```{r}

ActivityImputed<-Activity
ActivityImputed<-Activity %>% group_by(interval) %>% mutate(steps=replace(steps,is.na(steps), mean(steps,na.rm=TRUE)))
head(ActivityImputed)


```
### 4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r}

Step_Date_Imputed<-aggregate(steps~date, ActivityImputed, sum, na.rm=TRUE)
head(Step_Date_Imputed)

qplot(steps,data=Step_Date_Imputed,xlab='Total steps per day', ylab='Frequency using binwith 500',main='Histogram of total steps taken each day (Imputed)', binwidth=500)

```

The total numbers of steps for each day with imputed data are larger because all NAs are filled with the mean of the interval.

Calculate and report the mean and the median.
```{r}
mean(Step_Date_Imputed$steps)
median(Step_Date_Imputed$steps)
summary(Step_Date_Imputed)

```

## Are there differences in activity patterns between weekdays and weekends?
### 1.Create a new factor variable in the dataset with two levels – “weekday” and “weekend” 
```{r}
ActivityImputed$DayWeek<-weekdays(as.Date(ActivityImputed$date))
ActivityImputed$weekend<-ifelse(ActivityImputed$DayWeek %in% c('Saturday', 'Sunday'),'weekend', 'weekday')
head(ActivityImputed)
```

### 2. Make a panel plot containing a time series plot  of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
Step_interval_Imputed<-aggregate(ActivityImputed$steps, by=list(ActivityImputed$weekend,ActivityImputed$interval), FUN= "mean")
names(Step_interval_Imputed)<-c("weekend","interval","steps")
head(Step_interval_Imputed)
ggplot(Step_interval_Imputed, aes(x=interval, y=steps, color=weekend)) + geom_line()+
facet_grid(weekend ~.) + xlab("Interval") + ylab("Mean of Steps") +
    ggtitle("Comparison of Average Number of Steps in Each Interval Taken on Weekend or Weekday")
```

